WHITE WINE INVESTIGATION by Franz Hastrup-Nielsen \
date: May 15, 2018 \
output: html_document \
========================================================

```{r echo=FALSE, message=FALSE, warning=FALSE, packages}
# Notice that the parameter "echo" was set to FALSE for this code chunk. This
# prevents the code from displaying in the knitted HTML output. 
# The other parameters for "message" and "warning" should also be set to FALSE
# for other code chunks once verified that each plot comes out as expected.

library(ggplot2)
library(knitr)
library(dplyr)
library(GGally)
library(plyr)
library(RColorBrewer)
require(reshape2)
```

This report is going through the steps of Exploratory Data Analysis (EDA) using
a dataset that contains information about white wines and their associated
quality.  

```{r echo=FALSE, Load_the_Data}
# Change working directory here to run code
setwd(paste0("C:/Users/franz/Google Drev/Data Scientist Materiale",
      "/Data Analysis NanoDegree Udacity Part 2/Project 6 EDA - R",sep=""))
# Load the Data
pf <- read.csv('wineQualityWhites.csv')

```


# Univariate Plots Section

Input variables and short explanation: \
   1 - fixed.acidity              fixed acidity (tartaric acid - g / dm^3) \
   2 - volatile.acidity           volatile acidity (acetic acid - g / dm^3) \
   3 - citric.acid                citric acid (g / dm^3) \
   4 - res.sugar                  residual sugar (g / dm^3) \
   5 - chlorides                  chlorides (sodium chloride - g / dm^3 \
   6 - free.so2                   free sulfur dioxide (mg / dm^3) \ 
   7 - total.so2                  total sulfur dioxide (mg / dm^3) \
   8 - density                    density (g / cm^3) \
   9 - pH                         pH \
   10 - sulphates                 sulphates (potassium sulphate - g / dm3) \
   11 - alcohol                   alcohol (% by volume) \
   Output variable (based on sensory data):  \
   12 - quality                   quality (score between 0 and 10) \

```{r echo=FALSE}
dim(pf)
```
The dataset includes 4898 rows of wines and 12 columns assigned to each wine.

Here are the variables and a summary of the dataset to be investigated:

```{r echo=FALSE,message=FALSE}
pf <- plyr::rename(pf, c("residual.sugar"="res.sugar",
                   "free.sulfur.dioxide"="free.so2",
                   "total.sulfur.dioxide"="total.so2"))
names(pf)
head(pf)
summary(pf)
```


```{r echo=FALSE, Univariate_Plots}
qplot(fixed.acidity, data=pf, binwidth = 0.1)
qplot(y=pf$fixed.acidity, x= 1, geom = "boxplot")
summary(pf$fixed.acidity)
```

```{r echo=FALSE, Univariate_Plots2}
qplot(volatile.acidity, data=pf, binwidth = 0.01)
qplot(y=pf$volatile.acidity, x= 1, geom = "boxplot")
summary(pf$volatile.acidity)
```

Both the fixed and volatile acidity seems to be normally distributed with a 
positive skew, as a small tail forms towards the higher values. As the boxplot
shows, there seems to be a larger range of high values for both cases.


```{r echo=FALSE, Univariate_Plots3}
qplot(citric.acid, data=pf, binwidth = 0.02)
qplot(y=pf$citric.acid, x= 1, geom = "boxplot")
summary(pf$citric.acid)
```

Citric acid is normally distributed with a positive skew due to a longer tail
towards the higher values. Furthermore, it seems that two values on the right 
side is more prevalent close to 0.5 and 0.75. This could need further 
investigation. The boxplot shows the positive skew as the median is closer to
the 25% quantile.

```{r echo=FALSE, Univariate_Plots4}
qplot(res.sugar, data=pf, binwidth = 0.5)
qplot(y=pf$res.sugar, x= 1, geom = "boxplot")
summary(pf$res.sugar)
```

Residual sugar has a normal distribution with positive skew and some outliers
at very high values as indicated by the boxplot.

```{r echo=FALSE, Univariate_Plots5}
qplot(chlorides, data=pf, binwidth = 0.001)
qplot(y=pf$chlorides, x= 1, geom = "boxplot")
summary(pf$chlorides)
```

Chlorides behaves like the rest of the data with a normal distribution with a
positive skew. With a low binwidth the distribution could be displaying some 
bi-modal behavior

```{r echo=FALSE, Univariate_Plots6}
qplot(free.so2, data=pf, binwidth = 1)
qplot(y=pf$free.so2, x= 1, geom = "boxplot")
summary(pf$free.so2)
```

```{r echo=FALSE, Univariate_Plots7}
qplot(total.so2, data=pf, binwidth = 2)
qplot(y=pf$total.so2, x= 1, geom = "boxplot")
summary(pf$total.so2)
```

The free and total sulfur dioxide are both normally distributed with a
positive skew

```{r echo=FALSE, Univariate_Plots8}
qplot(density, data=pf, binwidth = 0.0005)
qplot(y=pf$density, x= 1, geom = "boxplot")
summary(pf$density)
```

```{r echo=FALSE, Univariate_Plots9}
qplot(pH, data=pf, binwidth = 0.01)
qplot(y=pf$pH, x= 1, geom = "boxplot")
summary(pf$pH)
```

```{r echo=FALSE, Univariate_Plots10}
qplot(sulphates, data=pf, binwidth = 0.01)
qplot(y=pf$sulphates, x= 1, geom = "boxplot")
summary(pf$sulphates)
```

Density, pH and sulphates all shows a normal distribution with a positive skew

```{r echo=FALSE, Univariate_Plots11}
qplot(alcohol, data=pf, binwidth = 0.1)
qplot(y=pf$alcohol, x= 1, geom = "boxplot")
summary(pf$alcohol)
```

The alcohol is normally distributed with a very positive skewed distribution.

```{r echo=FALSE, Univariate_Plots12}
qplot(quality, data=pf, binwidth = 1)
qplot(y=pf$quality, x= 1, geom = "boxplot")
summary(pf$quality)
```

The quality is normally distributed with a negative skew


# Univariate Analysis

### What is the structure of your dataset?\

**The dataset contains information about the quality of white wine. It has 4898 
observations with 12 different variables.**

### What is/are the main feature(s) of interest in your dataset? \

**The main features that will be investigated is how the different variables 
affect the quality, e.g. acidity, sugar, chlorides, sulphates and alcohol 
level.**

### Of the features you investigated, were there any unusual distributions? \

**Most of the data is found to have a normal distribution with a positive skew.
For several of the variables it seems that they include some outliers. These
could be removed if deemed necessary for this investigation. **

# Bivariate Plots Section

An overview of all the variables and their correlations are created:

```{r echo=FALSE, message=FALSE}
#ggpairs(pf[,2:12])
ggcorr(pf[,13:2],palette = "RdBu",label=FALSE,legend.position = "right")


```

Correlations that seems significant from this overview will be investigated
further below. The residual sugar seems to be correlated to many parameters and
will be investigated first. \
From previous analysis above, it was evident that residual sugar has one big 
outlier at 65.8, which will be removed from here on. Furthermore, free sulfur 
dioxide has one outlier at 289, which also will be removed. \



```{r echo=FALSE}
pf = subset(pf, res.sugar < 60)
pf = subset(pf, free.so2 < 200)
ggplot(aes(x=res.sugar,y=free.so2),data=pf) + 
  geom_point(alpha=0.1) +
  geom_smooth(color='red', method='lm')
#cor.test(pf$res.sugar, pf$free.so2, method = 'pearson')
pear = with(pf, cor(res.sugar, free.so2))
sprintf("Pearsons Correlation Coefficient: %f",pear)
```
An alpha value of 0.1 has been chosen to show the distribution of the data 
points. The free sulfur dioxide has a positive correlation of 0.32 with the 
residual sugar. There is a lot of wines with a very low amount of residual
sugar.

```{r echo=FALSE}
ggplot(aes(x=res.sugar,y=total.so2),data=pf) + 
  geom_point(alpha=0.1) +
  geom_smooth(color='red', method='lm')
#cor.test(pf$res.sugar, pf$total.so2, method = 'pearson')
pear = with(pf, cor(res.sugar, total.so2))
sprintf("Pearsons Correlation Coefficient: %f",pear)

```

The free and total sulfur dioxide behaves similarly when plotted against the 
residual sugar. For the total sulfur dioxide there is a positive correlation of
0.41 with the residual sugar. 

```{r echo=FALSE}
ggplot(aes(x=res.sugar,y=density),data=pf) + 
  geom_point(alpha=0.1) + 
  geom_smooth(color='red', method='lm')
#cor.test(pf$res.sugar, pf$density, method = 'pearson')
pear = with(pf, cor(res.sugar, density))
sprintf("Pearsons Correlation Coefficient: %f",pear)

```

The correlation between density and residual sugar is positive at 0.83, which is
a very high value and indicates a strong correlation. This is also evident from 
the plot. The datapoints for a low amount of residual sugar is less spread out 
for the density when compared to the sulfur dioxide.

```{r echo=FALSE}
ggplot(aes(x=res.sugar,y=alcohol),data=pf) +
  geom_point(alpha=0.1) + 
  geom_smooth(color='red', method='lm')
#cor.test(pf$res.sugar, pf$alcohol, method = 'pearson')
pear = with(pf, cor(res.sugar, alcohol))
sprintf("Pearsons Correlation Coefficient: %f",pear)

```

There is a negative correlation of -0.46 between alcohol and residual sugar.
There is a clear declining trend to the upper value of alcohol with increasing
residual sugar. \
\
All the plots above here, investigates the residual sugar and the significant 
correlations found. The two first plots are the free and the total sulfur 
dioxide. Persons correlation yields 0.32 and 0.41 for these two plots, which 
could indicate a moderate correlation between these and residual sugar. Both 
have a lot of datapoints with very little residual sugar, but otherwise there 
seems to be a positive correlation. \
The residual sugar and density seems to be well correlated, with a positive 
Pearson correlation coefficient of 0.83.\
Residual sugar and alcohol also seems to show some correlation with a negative 
Pearson correlation coefficient of -0.46. All of these datasets shows that a 
lot of wines have very little residual sugar, but a correlation seems to be 
prevalent for an increasing amount of residual sugar.

```{r echo=FALSE}
ggplot(aes(x=total.so2,y=free.so2),data=pf) +
  geom_point(alpha=0.1) +
  geom_smooth(color='red', method='lm')
#cor.test(pf$total.so2, pf$free.so2, method = 'pearson')
pear = with(pf, cor(total.so2, free.so2))
sprintf("Pearsons Correlation Coefficient: %f",pear)

```

There is a moderate to strong relationship between the free sulfur dioxide and 
the total sulfur dioxide with a positive correlation coefficient of 0.61. 

```{r echo=FALSE}

ggplot(aes(x=total.so2,y=density),data=pf) +
  geom_point(alpha=0.1) +
  geom_smooth(color='red', method='lm')
#cor.test(pf$total.so2, pf$density, method = 'pearson')
pear = with(pf, cor(total.so2, density))
sprintf("Pearsons Correlation Coefficient: %f",pear)


```

The total sulfur dioxide seems to be correlated with the density. The Pearsons 
correlation coefficient is 0.54, which indicates a moderate to good 
positive correlation.

```{r echo=FALSE}
ggplot(aes(x=pH,y=fixed.acidity),data=pf) + 
  geom_point(alpha=0.1) + 
  geom_smooth(color='red', method='lm')
#cor.test(pf$pH, pf$fixed.acidity, method = 'pearson')
pear = with(pf, cor(pH, fixed.acidity))
sprintf("Pearsons Correlation Coefficient: %f",pear)

```

pH and fixed acidity seems to be negatively correlated with a Pearsons 
correlation coefficient of -0.43. It makes sense to find a correlation between
these, as a low pH is a sign of high acidic content.

```{r echo=FALSE}
ggplot(aes(x=alcohol,y=total.so2),data=pf) +
  geom_point(alpha=0.1) + 
  geom_smooth(color='red', method='lm')
#cor.test(pf$alcohol, pf$total.so2, method = 'pearson')
pear = with(pf, cor(alcohol, total.so2))
sprintf("Pearsons Correlation Coefficient: %f",pear)

```

There is a negative relationship between alcohol and total sulfur dioxide of 
-0.45. There seems to be a decreasing scatter of total sulfur dioxide with 
increasing alcohol content.

```{r echo=FALSE}
ggplot(aes(x=alcohol,y=density),data=pf) +
  geom_point(alpha=0.1) + 
  geom_smooth(color='red', method='lm')
#cor.test(pf$alcohol, pf$density, method = 'pearson')
pear = with(pf, cor(alcohol, density))
sprintf("Pearsons Correlation Coefficient: %f",pear)
```

These plots investigate the correlations with alcohol. The residual sugar 
correlation was already investigated earlier with a Pearsons correlation 
coefficient of -0.46. Here the total sulfur dioxide has a negative Pearsons
correlation coefficient of -0.45. Looking at the plot, there is some scatter
but overall the correlation seems to be evident. \
Alcohol and density seems to be well correlated with a Pearsons correlation 
coefficient of -0.8. This was expected as the density of alcohol is less than 
water, such that an increase in alcohol will lower the overall density of the
wine.\
\

From the correlation overview made in the start of this section, only alcohol 
seemed to have some significant correlation with quality and to some extent 
density as well. These will be investigated here. 
```{r echo=FALSE}
ggplot(pf,aes(factor(quality),alcohol))+
  geom_boxplot(aes(color=factor(quality)))+
  geom_jitter(alpha=0.05, aes(color=factor(quality)),
              position=position_dodge(width=0.5))

```

There is a tendency that the higher alcohol content wines have a higher quality. 
However, for the lowest quality, the alcohol content seems to be increasing a
bit as well. Wine with a quality of 5 seems to be skewed.

```{r echo=FALSE}
ggplot(pf,aes(factor(quality),density))+
  geom_boxplot(aes(color=factor(quality)))+
  geom_jitter(alpha=0.05,aes(color=factor(quality)),
              position=position_dodge(width=0.5))
```

The density and quality shows a reverse trend to what was seen in the previous
plot with alcohol. As alcohol and density are so closely related, this trend 
is given due to the physical relationship between alcohol content and density.\
\
Below here it will be investigated if other variables could have a significant
influence on the quality.

```{r echo=FALSE, Bivariate_Plots}
ggplot(pf,aes(factor(quality),fixed.acidity))+
  geom_boxplot(aes(color=factor(quality)))+
  geom_jitter(alpha=0.05,aes(color=factor(quality)),
              position=position_dodge(width=0.5))
```

It seems that for medium to high quality the fixed acidity is close to constant.
Only the lowe quality wines have an increase in fixed acidity and a large spread
of data.

```{r echo=FALSE}
ggplot(pf,aes(factor(quality),volatile.acidity))+
  geom_boxplot(aes(color=factor(quality)))+
  geom_jitter(alpha=0.05,aes(color=factor(quality)),
              position=position_dodge(width=0.5))
```

The volatile acidity is high for low and high quality wines. The lowest volatice
acidity is found for the medium quality white wines.

```{r echo=FALSE}
ggplot(pf,aes(factor(quality),res.sugar))+
  geom_boxplot(aes(color=factor(quality)))+
  geom_jitter(alpha=0.05,aes(color=factor(quality)),
              position=position_dodge(width=0.5))
```

The lowest value of residual sugar seems to be rather constant for all qualities
but the highest values varies with quality. Most of the datasets are skewed.

```{r echo=FALSE}
ggplot(pf,aes(factor(quality),citric.acid))+
  geom_boxplot(aes(color=factor(quality)))+
  geom_jitter(alpha=0.05,aes(color=factor(quality)),
              position=position_dodge(width=0.5))
```

The quality does not seem to vary a lot with the citric acidic values. The 
highest quality wines seems to have more citric acid.

```{r echo=FALSE}
ggplot(pf,aes(factor(quality),pH))+
  geom_boxplot(aes(color=factor(quality)))+
  geom_jitter(alpha=0.05,aes(color=factor(quality)),
              position=position_dodge(width=0.5))
```

It could seem that an increasing pH could mean a higher quality wine, however
the low quality wines also has a high pH. pH alone is not an indicator of 
quality.

```{r echo=FALSE}
ggplot(pf,aes(factor(quality),chlorides))+
  geom_boxplot(aes(color=factor(quality)))+
  geom_jitter(alpha=0.05,aes(color=factor(quality)),
              position=position_dodge(width=0.5))
```

For the cholride content the quartiles in the boxplot are very narrow in range
with a large range of outliers. The median trend is for high quality wines to
have a low content of chlorides and the outliers seems to narrow in as well.

```{r echo=FALSE}
ggplot(pf,aes(factor(quality),sulphates))+
  geom_boxplot(aes(color=factor(quality)))+
  geom_jitter(alpha=0.05,aes(color=factor(quality)),
              position=position_dodge(width=0.5))
```

The median for sulphatesshows a nearly constant trend and does not seem to have
a major impact on the quality of the wine. 

```{r echo=FALSE}
ggplot(pf,aes(factor(quality),total.so2))+
  geom_boxplot(aes(color=factor(quality)))+
  geom_jitter(alpha=0.05,aes(color=factor(quality)),
              position=position_dodge(width=0.5))
```

There seems to be a trend of decreasing total sulfur dioxide for higher quality
wines.

```{r echo=FALSE}
ggplot(pf,aes(factor(quality),free.so2))+
  geom_boxplot(aes(color=factor(quality)))+
  geom_jitter(alpha=0.05,aes(color=factor(quality)),
              position=position_dodge(width=0.5))
```

Both low and high quality wines have a low amount of free sulfur dioxide. This
makes it difficult to identify the wine quality from this parameter.

# Bivariate Analysis

### Talk about some of the relationships you observed in this part of the \
investigation. How did the feature(s) of interest vary with other features in \
the dataset?

**This dataset seems to only contain a limited amount of very high and very low 
quality wines but many more in the middle. This results in a lower spread of the
variables when investigating the low and high quality wines, whereas a larger 
spread in the data is seen for the ratings in between. \
For some of the variables, it seems that they are converging on a very limited 
range for a high quality wine, e.g. when looking at the pH. This small range of 
high quality might not represent the truth due to the limited amount of high 
quality wines in the dataset. \
\
Most of the datasets plotted against quality, showed either an increasing or 
decreasing  trend to the best quality. However in most of them, the lowest 
quality wines usually also showed the same trend as the high quality wines, 
which makes it difficult to properly identify the best wines from these physical
parameters.**


### Did you observe any interesting relationships between the other features \
(not the main feature(s) of interest)?

**There seems to be a strong relationship between alcohol and both residual 
sugar and density. The density was expected, as alcohol has a lower density than 
water, hence increasing the alcohol will decrease the wines density.\
Another expected correlation was between citric acid and the pH level, as pH 
will decrease when more acid is present.**

### What was the strongest relationship you found?

**Two strong relationships were found. The density and alcohol content has a 
Pearson correlation coefficient of -0.80 and residual sugar and density has a 
Pearson correlation coefficient of 0.83.\
Density is clearly influenced by both the alcohol content and residual sugar. 
Increasing the alcohol content will lower the density, as it is lighter than 
water. Increasing the residual sugar increases the density.**

# Multivariate Plots Section

To ease the comparison between the qualities of wine, they will be collected 
into three groups:\
Low quality: <= 5 \
Medium quality: 6-7 \
High quality: => 8 \

```{r echo=FALSE,message=FALSE, warning=FALSE, Multivariate_Plots}
pf = na.omit(pf)
df3 <- pf %>% 
  mutate(quality.short = if_else(quality < 6 ,
                                 'Low quality',ifelse(quality > 7,
                                            'High quality', 'Medium quality')))

df3$quality.short <- factor(df3$quality.short, levels =c("High quality",
                                                         "Medium quality",
                                                         "Low quality"))

ggplot(df3,aes(x=res.sugar, y=alcohol, color=as.factor(quality.short), 
               shape=as.factor(quality.short))) +
  geom_point(alpha=0.75) + 
  geom_smooth(method=lm, se=FALSE, fullrange=TRUE,size=2)+ 
  xlim(0,35)+
  scale_shape_manual(values=c('Low','Med','High')) + 
  theme(legend.position="right")

ggplot(df3, aes(x=res.sugar, y=density, color=as.factor(quality.short), 
                shape=as.factor(quality.short))) +
  geom_point(alpha=0.25) + 
  geom_smooth(method=lm, se=FALSE, fullrange=TRUE,size=2)+ 
  xlim(0,35)+
  ylim(0.985,1.01) +
  scale_shape_manual(values=c('Low','Med','High'))  +
  theme(legend.position="right")

```

The first plot, shows the relationship between residual sugar and alcohol. The 
three lines are showing the low, medium and high quality wines. It is quite 
evident, that the higher alcohol content, the higher quality. All 3 lines almost
intersects in the same place, but this is outside the data range and does not 
seem to be valid. \
The second plot, shows the relationship between residual sugar and density The 
three lines are showing the low, medium and high quality wines. For this plot, 
the order of the lines have been reversed. High quality wines have a low density
and low quality wines have a high density - this is again related to alcohol. 

```{r echo=FALSE,message=FALSE, warning=FALSE}
#df2 <- read.csv('wineQualityWhites.csv', header=T)
df2 <- df3[,c(2,5,12,14)]
df2.m <- melt(df2, id.vars = "quality.short")
df2.m$quality.short <- factor(df2.m$quality.short, levels =c("High quality",
                                                             "Medium quality",
                                                             "Low quality"))
ggplot(data = df2.m, aes(x=variable, y=value)) + 
  geom_boxplot(aes(fill=quality.short)) + ylim(0,35) +
scale_fill_manual(name="Quality", values = c("#0033FF","#0072B2","#56B4E9"))
```

The median for ficed acidity seems to increase slightly for decreasing quality, 
whereas it is opposite for alcohol. For residual sugar, the level is almost the 
same for high and medium quality wines, where low quality wines have a much 
higher median.

```{r echo=FALSE,message=FALSE, warning=FALSE}
df4 <- df3[,c(7,8,14)]
df4.m <- melt(df4, id.vars = "quality.short")
df4.m$quality.short <- factor(df4.m$quality.short, levels =c("High quality",
                                                             "Medium quality",
                                                             "Low quality"))
ggplot(data = df4.m, aes(x=variable, y=value)) + 
  geom_boxplot(aes(fill=quality.short)) +
scale_fill_manual(name="Quality", values = c("#0033FF","#0072B2","#56B4E9"))
```

The free sulfur dioxide content does not seem to influence the quality, but the 
lower total sulfur dioxide the better quality.

```{r echo=FALSE,message=FALSE, warning=FALSE}
df5 <- df3[,c(3,4,11,14)]
df5.m <- melt(df5, id.vars = "quality.short")
df5.m$quality.short <- factor(df5.m$quality.short, levels =c("High quality",
                                                             "Medium quality",
                                                             "Low quality"))
ggplot(data = df5.m, aes(x=variable, y=value)) + 
  geom_boxplot(aes(fill=quality.short)) +
scale_fill_manual(name="Quality", values = c("#0033FF","#0072B2","#56B4E9"))
```

Acidity, citric acid and sulphates does not seem to have a large influence on 
the quality of the wines as the medians are almost constant across the
qualities.

```{r echo=FALSE,message=FALSE, warning=FALSE}
df6 <- df3[,c(6,14)]
df6.m <- melt(df6, id.vars = "quality.short")
df6.m$quality.short <- factor(df6.m$quality.short, levels =c("High quality",
                                                             "Medium quality",
                                                             "Low quality"))
ggplot(data = df6.m, aes(x=variable, y=value)) + 
  geom_boxplot(aes(fill=quality.short)) +
scale_fill_manual(name="Quality", values = c("#0033FF","#0072B2","#56B4E9"))
```

It could seem that the for lower chlorides you get a the better quality wine.

```{r echo=FALSE,message=FALSE, warning=FALSE}
df7 <- df3[,c(9,14)]
df7.m <- melt(df7, id.vars = "quality.short")
df7.m$quality.short <- factor(df7.m$quality.short, levels =c("High quality",
                                                             "Medium quality",
                                                             "Low quality"))
ggplot(data = df7.m, aes(x=variable, y=value)) + 
  geom_boxplot(aes(fill=quality.short)) +
scale_fill_manual(name="Quality", values = c("#0033FF","#0072B2","#56B4E9"))
```

The lower density the higher quality. This is the reverse from alcohol as seen
earlier and is expected due to the physical dependency between the two.
 
```{r echo=FALSE,message=FALSE, warning=FALSE}
df8 <- df3[,c(10,14)]
df8.m <- melt(df8, id.vars = "quality.short")
df8.m$quality.short <- factor(df8.m$quality.short, levels =c("High quality",
                                                             "Medium quality",
                                                             "Low quality"))
ggplot(data = df8.m, aes(x=variable, y=value)) + 
  geom_boxplot(aes(fill=quality.short))+
scale_fill_manual(name="Quality", values = c("#0033FF","#0072B2","#56B4E9"))
```

This plot indicated that for a higher pH you get a better quality white wine.


# Multivariate Analysis

### Talk about some of the relationships you observed in this part of the \
investigation. Were there features that strengthened each other in terms of \
looking at your feature(s) of interest?

**Some of the relationships between the different variables and quality, that was
investigated in the bivariate section, has become more clear. E.g. the alcohol 
vs sectioned quality only indicates higher quality with increasing
alcohol, although the plot in the bivariate section showed a slight increase in
alcohol for low quality wines. This is due to the sectioning of quality into 
low, medium and high quality. \
This sectioning could remove some trends, e.g. the tendency that was seen in
bivariate analysis with high and low quality wines showing the same behaviors.**

### Were there any interesting or surprising interactions between features?

**I would have assumed that citric acidity, volatile acidity and sulphates would 
have had a larger influence on the perception of the quality of white wine. From
this investigation, it does not seem that these variables have a significant 
influence and other parameters are of much more interest.**

------

# Final Plots and Summary


### Plot One
```{r echo=FALSE, Plot_One}

qplot(pH, data=pf, binwidth = 0.01) + ggtitle("Distribution of pH values") + 
  ylab("Count") +
  xlab("pH")
me = mean(pf$pH)
med = median(pf$pH)
sprintf("The pH median is %.1f and the mean is %.1f",med,me)

```

### Description One

**The pH variable is a nice example of a normal distribution with the mean
equal to the median indicating no skew.**

### Plot Two
```{r echo=FALSE, Plot_Two}
ggplot(aes(x=res.sugar,y=density),data=pf) + 
  geom_point(alpha=0.1) + 
  geom_smooth(color='red', method='lm') +
    ggtitle("Density as a function of residual sugar") + 
  xlab("Residual sugar Content [g/dm^3]") +
  ylab("Density [g / cm^3]")
pear = with(pf, cor(res.sugar, density))
sprintf("Pearsons Correlation Coefficient: %f",pear)

```

### Description Two

**There is a strong positive correlation between density and the residual sugar.
This makes sense as increasing the sugar content of the wine will increase the
density.**

### Plot Three
```{r echo=FALSE, Plot_Three}
ggplot(pf,aes(factor(quality),alcohol))+
  geom_boxplot(aes(color=factor(quality)))+
  geom_jitter(alpha=0.05,aes(color=factor(quality)),
              position=position_dodge(width=0.5)) +
      ggtitle("Alcohol content as a function of quality") + 
  ylab("Alcohol Content [% by vol]") +
  xlab("Quality")

```

### Description Three
**The data seems to indicate that you should look for a high alcohol content and
preferably above 12% in order to get a nice white wine. Interestingly you could
be tricked when choosing a white wine with an alcohol content of ~10.5%, as it 
could have a quality rating of both 3 and 6. **

------

# Reflection

**It is surprising that citric acid, sulphates and volatile acidity does not seem 
to have a big influence on the quality of white wines, whereas alcohol and 
density seems to be the most important factors. Density is closely correlated 
to alcohol content and residual sugar as these two impacts the density of the 
wine. From all the parameters a high alcohol content is the key indicator for a 
good quality white wine. \
This evaluation builds on physical parameters of white wine and some quality 
value. According to the dataset, the quality parameter has been established 
using "sensory data". Wine taste and quality are very subjective and will most 
likely vary from person to person, which is why this quality parameter can be 
hard to trust. To test some of the conclusions from this dataset a blind tasting 
with a large group of people could be used to verify the conclusions.\
\
One of the struggles with this dataset has been to identify the correlations 
between the variables and finding the greatest indicators of a high quality 
white wine. After identifying relationships and correlations I have spent a 
great deal of time contemplating how to show the data and which plots and R
functions that could be utilized.
**
